Quantum Hoare Type Theory

### 2020-11-09 ms-paper/ms-paper.tex
Remove unneeded page breaks

\paragraph{Verification}
We would like to describe how QHTT lets us verify the correctness of the above program based on the specification provided as the type of the function. For illustration, we show the propositions annotated as comments as we step through the \texttt{testBell00} program below.
\newpage
\begin{lstlisting}[language=QHaskell]
testBell00 : {\top} (a, b) : (Bool o* Bool) {\top /\ a = b}
              -- \top
         = do qa <= init false;
              -- qa =q |0\>, class(qa)
              apply H to qa;
              -- qa =q |+\>, uniform(qa)
              qb <= init false;
              -- qa =q |+\>, uniform(qa), qb =q |0\>, class(qb)
              apply CNOT to (qa, qb);
              -- (qa, qb) =q |++\>  -- can use a weaker but more general `EPR' here
              -- /\ uniform(qa, qb) -- this won't apply then
              a = meas qa;
              --   (qa, e1) =q |00\>, (qb, e2) =q |00\>, a = false
              -- \/ (qa, e1) =q |11\>, (qb, e2) =q |11\>, a = true
              b = meas qb;
              --   (qa, e1) =q |00\>, (qb, e2) =q |00\>, a = false, b = false
              -- \/ (qa, e1) =q |11\>, (qb, e2) =q |11\>, a = true,  b = true
              return (a, b)
\end{lstlisting}

@@

\section{Implementation}
We are implementing QHTT in the F* programming language.

One issue with the original formulation of Hoare Type Theory is that using Hoare triples does not directly lead to a verification condition generation algorithm. This results in requiring a complicated bidirectional typing system for implementation. If we use proposition transformers as demonstrated by the use of Dijkstra monads in F*, we can avoid these issues \cite{swamy2013verifying}.

The current steps for the implementation are:

\begin{itemize}
	\item Design a memory model for representing quantum state, along with elementary operations such as read, write, lookup, etc
	\item Expose a new Quantum effect in F* by using the Dijkstra Monads for Free technique \cite{dm4free2017}.
	\item Port our language to work well with F* syntax.
\end{itemize}

This will necessarily take us away from the bidirectional typing approach used in the previous section. But the end result for the user would be the same. Indeed, F* system is an evolution of the HTT approach that comes with inherent support for effects and automation using an SMT solver.

%\section{Evaluation}

%\subsection{More Examples}
%\textit{add at least one example with scaling knob}

%\subsection{Circuit-complexity Considerations}

>>>>

\paragraph{Verification}
We would like to describe how QHTT lets us verify the correctness of the above program based on the specification provided as the type of the function. For illustration, we show the propositions annotated as comments as we step through the \texttt{testBell00} program below.

\begin{lstlisting}[language=QHaskell]
testBell00 : {\top} (a, b) : (Bool o* Bool) {\top /\ a = b}
              -- \top
         = do qa <= init false;
              -- qa =q |0\>, class(qa)
              apply H to qa;
              -- qa =q |+\>, uniform(qa)
              qb <= init false;
              -- qa =q |+\>, uniform(qa), qb =q |0\>, class(qb)
              apply CNOT to (qa, qb);
              -- (qa, qb) =q |++\>  -- can use a weaker but more general `EPR' here
              -- /\ uniform(qa, qb) -- this won't apply then
              a = meas qa;
              --   (qa, e1) =q |00\>, (qb, e2) =q |00\>, a = false
              -- \/ (qa, e1) =q |11\>, (qb, e2) =q |11\>, a = true
              b = meas qb;
              --   (qa, e1) =q |00\>, (qb, e2) =q |00\>, a = false, b = false
              -- \/ (qa, e1) =q |11\>, (qb, e2) =q |11\>, a = true,  b = true
              return (a, b)
\end{lstlisting}

@@
